---
title: "206 Lab 5"
author: "Wes"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####Lab 5 objectives 
-Learn to parse dates with "lubridate" 
-Get counts of obervations with count()
-Uncount() rows 
-One and two-sample t-test 
-Create a heatmap with geom_tile()

```{r, include = FALSE}
#Read in data and attach packages
library(tidyverse)
library(janitor)
library(here)

lobster_abundance <-read.csv(here::here("data","lobster_abundance.csv"),na = "-99999") %>% 
  clean_names() 
  
```

Use tidyr::uncount()  function to convert our lobster data from frequency format to case format 

```{r}
lobster_tidy <- lobster_abundance %>%  
  uncount(lobster_count)
```

#### Exploratory data visualization 

We're only going to consider "site" as our variable of interest 

```{r}
ggplot(lobster_tidy, aes(x= site, y= size_mm))+ 
  geom_jitter(aes(color= site), 
              alpha = .5, 
              width = .2)

#Histogram 
ggplot(lobster_tidy, aes(x= size_mm))+ 
  geom_histogram(aes(fill= site))+ 
  facet_wrap(~site, scales= "free") 

ggplot(lobster_tidy, aes(sample = size_mm))+ 
  geom_qq() + 
  facet_wrap(~site)
```
#### Convert the "date" column to class "Date" 

We'll use the lubridate package to convert to date format and then to help us easily parse month and year 

```{r}
lobster_date <- lobster_tidy %>%  
  mutate(date_new = lubridate::mdy(date))
```
Now lets parse year and month using lubridate::month() and lubridate::year()
```{r}
lobster_parse_date <- lobster_date %>%  
  mutate(obs_month = lubridate::month(date_new, label= TRUE), 
obs_year = lubridate::year(date_new))

```
Now let's find counts of observed lobsters based on different hierarchical groupings: 

First let's count lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>%  
  count(obs_year, obs_month) 

lobster_y <- lobster_parse_date %>%  
  count(obs_year) 

lobster_site <- lobster_parse_date %>%  
  count(site)

```

If we want to create a summary table that contains statistics other than counts by group, its easier to use group_by() +n()

```{r}
lobster_summary <- lobster_parse_date %>%  
  group_by(site) %>%  
  summarise(mean_size = mean(size_mm, na.rm= TRUE), 
            sd_size = sd(size_mm, na.rm = TRUE), 
                         lobster_number = n()) 

```

#### Find confidence intervals 

Us t.test() function to find confidence intervals for one sample and perform t-test to compare means of two samples

```{r}
ivee_lobster <- lobster_tidy %>%  
  filter(site == "IVEE") %>%  #== means matches with...
  pull(size_mm) 

t.test(ivee_lobster)
```

##### Two sample t-test to compare means 

Is there a significant difference in lobster lengths at Naples and Mohawk reefs? 

We have done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate. 

```{r}
napl_sample <- lobster_tidy %>%  
  filter(site== "NAPL") %>%  
  pull(size_mm)

mohk_sample <- lobster_tidy %>%  
  filter(site == "MOHK") %>% 
  pull(size_mm)

mm_ttest <- t.test(mohk_sample, napl_sample)
mm_ttest
```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p< 0.001, alpha = 0.05) 

```{r}
lobster_mn <- lobster_tidy %>%  
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn) 

mn_ttest2
mn_ttest2$p.value 
mn_ttest2$statistic
```
Here is the p-value: `r mn_ttest2$p.value`  
Here is my t-stastic: `r mn_ttest2$statistic` 

### Heatmaps: 

```{r}
lobster_ys <- lobster_parse_date %>%  
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x= obs_year, 
                              y= site))+ 
  geom_tile(aes(fill = n))
```

